---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)

setwd("/Volumes/Elements/D-Drive/University of Maryland/Classes/Fall 2022/Data Journalism/Github Repository/Data_Journalism_2022_Fall/labs/lab_02/")

```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?

A1. A majority of contributions are included in the amount category of $10 and $25 and the second greatest contribution amount are donations which are less than $10. It shows that people are not giving an incredibly large sum of money. 

```{r}
wv_winred_contribs_category <-wv_winred_contribs %>% 
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~"Between $10 and $25",
        amount > 25 & amount <= 50 ~ "Between $26 and $50",
        amount > 50 & amount <= 100 ~ "Between $51 and $100",
        amount > 100 & amount <= 500 ~ "Between $101 and $500",
        amount > 500 ~ "Greater than $500"
      )) %>% 
  group_by(amount_category) %>% 
  summarize(total_contributions = n())

```

Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  

A2. Rev Dr. John Almond made the most donations of under $10. He made 293 donations of under $10.

```{r}
wv_winred_contribs %>% 
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~"Between $10 and $25",
        amount > 25 & amount <= 50 ~ "Between $26 and $50",
        amount > 50 & amount <= 100 ~ "Between $51 and $100",
        amount > 100 & amount <= 500 ~ "Between $101 and $500",
        amount > 500 ~ "Greater than $500"
      )) %>% 
filter(amount < 10) %>% 
group_by(last_name, first_name, address_one) %>% 
  summarize(total_contributions = n()) %>% 
    arrange(desc(total_contributions))
```

Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.

A3. One question I have is what connection Dr. John Almond has to the political candidates and if the candidates have any connection to him in any way. 

Q4. Which month saw the most contributions?

A4. June saw the most contributions with 4,636. 

```{r}
wv_winred_contribs %>% 
 mutate(month = floor_date(date, "month")) %>%
  group_by(month) %>%
  summarize(total_contributions = n()) %>% 
  arrange(desc(total_contributions))
```

Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.

A5. There were 6 donations of exactly 2,900 and five of the donors were from West Virginia.

```{r}
wv_winred_contribs %>% 
filter(amount == 2900)
```
