---
title: "lab_03"
author: "Derek Willis"
date: "8/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)
# Load janitor here
install.packages('janitor')

setwd("/Volumes/Elements/D-Drive/University of Maryland/Classes/Fall 2022/Data Journalism/Github Repository/Data_Journalism_2022_Fall/labs/lab_03/")

```

## Load data
```{r}
# Write code to load MD slice of FEC House candidate contribution data and call it md_contribs.
# The data is called lab_03.rds, and is in the data folder alongside lab_03.Rmd (this data notebook)

 lab_03 <- read_rds("data/lab_03.rds")

```

## Answer questions

Q1: Which seven cities in Maryland had the largest _number_ of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1: The top seven cities are Baltimore (6726), Bethesda (5868), Silver Spring (4659), Chevy Chase (3260), Potomac (2301), Rockville (2213) and Annapolis (2017. 
```{r}
## Cleaning up data functions
## find what was spelled wrong...
lab_03 %>% 
  group_by(contributor_city) %>% 
  filter(str_detect(contributor_city, "^BALIT")) %>%
  count()

lab_03 %>% 
  group_by(contributor_city) %>% 
  filter(str_detect(contributor_city, "^BET")) %>% 
    count()
  
lab_03 %>% 
  group_by(contributor_city) %>% 
  filter(str_detect(contributor_city, "^SIL")) %>%
  count()

lab_03 %>%
  group_by(contributor_city) %>% 
  filter(str_detect(contributor_city, "^AN")) %>% 
  count()
  
lab_03 %>% 
  group_by(contributor_city) %>% 
  filter(str_detect(contributor_city, "^CHEV")) %>% 
  count()  

lab_03 %>% 
  group_by(contributor_city) %>% 
  filter(str_detect(contributor_city, "^POT")) %>%
  count()
  
 lab_03 %>% 
  group_by(contributor_city) %>% 
  filter(str_detect(contributor_city, "^ROCK")) %>%
  count()

 
lab_03 <- lab_03 %>% 
mutate(contributor_city = case_when(contributor_city == "ANRETIREDPOLIS," ~ "ANNAPOLIS",
                                    contributor_city == "ANNAPOLID" ~ "ANNAPOLIS",
                                    contributor_city == "ANNAPOLIS JUNCTION" ~ "ANNAPOLIS",
                                    contributor_city == "ANNNAPOLIS" ~ "ANNAPOLIS",
                                    contributor_city == "BETHESA" ~ "BETHESDA",
                                    contributor_city == "BETHSDA" ~ "BETHESDA",
                                    contributor_city == "BETTERTON" ~ "BETHESDA",
                                    contributor_city == "BETHESA" ~ "BETHESDA",
                                    contributor_city == "SILVER SPRING, MD" ~ "SILVER SPRING",
                                    contributor_city == "SILVER SPRINGS" ~ "SILVER SPRING",
                                    contributor_city == "SLIVER SPRING" ~ "SILVER SPRING",
                                    contributor_city == "POTAMIC" ~ "POTOMAC",
                                    contributor_city == "BALITMORE" ~ "BALTIMORE",
                                    TRUE ~ contributor_city
                                    ))

lab_03 %>% 
  group_by(contributor_city) %>% 
  count() %>% 
  arrange(desc(n))

new_city<- lab_03

                      
```

Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). Where is that zip code and what kind of area is it? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2: 20815 is chevy chase Maryland. It is a wealthy area. Jamie Raskin for Congress received the highest amount of contributions from zip code 20815. Two of the five committees are from Maryland, Jamie Raskin for Congressa and Heather Miezur are both from Maryland.
```{r}

##Mutating zip so that it is 5 digits
new_city <- new_city %>%
  mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))

##What zip code has the highest amount of contributions?
new_city %>% 
group_by(contributor_zip) %>% 
summarise(total_for_zip = sum(contribution_receipt_amount)) %>% 
arrange(desc(total_for_zip))

##committes that got most money from 20815
new_city %>% 
group_by(committee_name) %>% 
filter(contributor_zip == "20815") %>% 
summarise(total_for_committee = sum(contribution_receipt_amount)) %>% 
arrange(desc(total_for_committee))
  


# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
# Second, use the top ZIP code in the dataframe you just created and create a new dataframe that shows the committee that got the most money from that ZIP code.

```

Q3: Let's look at top individual donors to House campaigns. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. One of the top few donors has a very strong connection to the U.S. House - who is that and what does he do? How does his donation pattern differ from the second-largest donor, and what can you say about that second-largest donor?

Feel free to add any interesting "color" (little interesting details) you find in your research.  

A3: David Trone is a house of reps member. He is a democrat whose donations typically total 2,900. He donates to many other congressional democratic campaigns. The second largest donor is Juana Mason who donated over 3,000 times to various campaigns and varrying amounts. I also noticed that Trone spent a large sum of money a few amount of times and Mason donated smaller amounts but donated many more times.

```{r}
## Finding the highest amount of contributions and organizing from most to least
new_city %>% 
group_by(contributor_name, contributor_city) %>%
summarise(count_contribs = n(), total_amount = sum(contribution_receipt_amount)) %>% 
arrange(desc(total_amount))


## This shows the second highest donor.
new_city %>%
group_by(committee_name) %>%
filter(contributor_name == "MASON, JUANITA",
       contributor_city == "BALTIMORE") %>%
count() %>%
arrange(desc(n))

## Finding out which committees Mason donated the most money to
new_city %>%
group_by(committee_name) %>%
filter(contributor_name == "MASON, JUANITA",
       contributor_city == "BALTIMORE") %>%
summarise(total = sum(contribution_receipt_amount)) %>%
arrange(desc(total))

new_city %>%
  group_by(contributor_name) %>%
  count() %>%
  arrange(desc(n))

#NOTE: I was trying to figure out max and min for Juanita's donations but i kept getting error messages.

# Hint
# First, create a dataframe of top individual contributors
# Second, use information from that dataframe to examine individual donations from the original md_contribs dataframe.

```
