---
title: "lab_06"
author: "derek willis"
date: "10/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Tabula

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the libraries you will need

options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

setwd("/Volumes/Elements/D-Drive/University of Maryland/Classes/Fall 2022/Data Journalism/Github Repository/Data_Journalism_2022_Fall/labs/lab_06/")

```

## Get Our PDF

We'll be working with the [Maryland monthly voter registration update](https://elections.maryland.gov/pdf/vrar/2022_09.pdf) from the State Board of Elections. You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to write a couple of paragraphs that summarize the changes in voter registration in Maryland during the month of September.

## Setup Tabula

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds.

This PDF has multiple possible tables to extract. We're going to make four dataframes: new registrations by party, removals by party, changes in registration and current registrations by county. You will need to draw boxes around four sections and export four CSV files that you will load into R. In Tabula, draw a box around the borders of those tables and click the "Preview & Export Extracted Data" button for each one. The export should look pretty clean.

Export each CSV (each one should be called `tabula-2022-09.csv` by default, so you should rename them for clarity) to your lab_06/data folder.

From there, you will need to read in and clean up the data, and in particular the headers. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)).

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers should not contain spaces and they should have meaningful names, not "x1" or something similar. The current registrations dataframe *MUST* include the county name, but should not include the columns under `Changes` and `Party Affiliation From`. How you do that is up to you (and could involve something outside R), but you can use select() with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns.

```{r}
# Party registrations

new_party_registration_2022 <- read_csv("data/new_party_registration_2022.csv")

new_party_registration_2022 <- new_party_registration_2022 %>% 
clean_names()

new_party_registration_2022 <- new_party_registration_2022 %>% 
rename(county = x1)




```

```{r}
# Removals
### Need to clean up the blank columns by removing the NAs

removals_2022 <- read_csv("data/removals_2022.csv")

removals_2022 <- removals_2022 %>% 
clean_names()

removals_2022 <- removals_2022 %>% 
rename(removal_method = x1)

removals_2022 <- removals_2022 %>% 
filter(!is.na(removal_method))


```

```{r}
# Changes

party_changes_2022 <- read_csv("data/party_changes_2022.csv")

party_changes_2022 <- party_changes_2022 %>% 
clean_names()

party_changes_2022 <- party_changes_2022 %>% 
rename(county = x2) %>%
select(-x1, -x3, -x6, -x15) 


```

```{r}
# Current registrations

current_registration_2022 <- read_csv("data/current_registration_2022.csv")

current_registration_2022 <- current_registration_2022 %>% 
clean_names() 

current_registration_2022 <- current_registration_2022 %>% 
select(-dem_4, -rep_5, -grn_6, -lib_7, -wcp_8, -oth_10, -total_11)
### try to negative select the left part of the table to I am able to have the county column with the right side of the data. 

current_registration_2022 <- current_registration_2022 %>% 
rename(county = x1, active = x21, conf_mailing = inactivemailingconf) %>% 
select(-unaf_9)


```

## Rinse, Repeat: September 2020 data

Do all of that again, for the [September 2020 voter registration update](https://elections.maryland.gov/pdf/vrar/2020_09.pdf), so we can compare the two. If you've done well, this is easy: just do the same steps in Tabula, export the files and use the same code to read in and clean up the data, *just remember to give these dataframes different names*, although similar enough so that you know how to compare September 2022 to September 2020.

To answer the following questions, you'll need to calculate percentages and do other calculations. Whether you do that by combining or joining dataframes or you write code to run separately on both is up to you.

```{r}
# Party registrations

new_party_registration_2020 <- read_csv("data/new_party_registration_2020.csv")

new_party_registration_2020 <- new_party_registration_2020 %>% 
clean_names()

new_party_registration_2020 <- new_party_registration_2020 %>% 
rename(county = x1)

```

```{r}
# Removals

removals_2020 <- read_csv("data/removals_2020.csv")

removals_2020 %>% 
clean_names() %>% 
rename(removal_method = x1)

removals_2020 <- removals_2020 %>% 
filter(!is.na(removal_method))



```

```{r}
# Changes

party_changes_2020 <- read_csv("data/party_changes_2020.csv")

party_changes_2020 <- party_changes_2020 %>% 
clean_names() %>% 
rename(county = x1)
  
party_changes_2020 <- party_changes_2020 %>% 
mutate(county = case_when(county == "PR. GEORGE'S" ~ "PRINCE GEORGE'S", 
                        TRUE ~ county))


```

```{r}
# Current registrations

current_registration_2020 <- read_csv("data/current_registration_2020.csv")

current_registration_2020 <- current_registration_2020 %>% 
clean_names()

current_registration_2020 <- current_registration_2020 %>% 
rename(county = x2)

current_registration_2020 <- current_registration_2020 %>% 
select(-x1, -x3, -x6, -dem_7, -rep_8, -grn_9, -lib_10, -wcp_11, -bar_12, -unaf_13, -oth_14, -total_15, -x16)

current_registration_2020 <- current_registration_2020 %>% 
select(-x26, -x29)

current_registration_2020 <- current_registration_2020 %>% 
mutate(county = case_when(county == "PR. GEORGE'S" ~ "PRINCE GEORGE'S", 
                        TRUE ~ county))

```


## Answer questions

Q1. Calculate the percentage of registered Democrats, Republicans and unaffiliated voters for the entire state in September 2022 and September 2020, and then calculate the percentage change for those three party categories (the formula for percentage change is easy to remember: (New-Old)/Old). How would you describe the changes in these partisan categories in a story? What's the most interesting thing about these changes that would be the lede of that story?

A1. Dems 2022% = 54.1%, Dems in 2020 = 55%, percent change is a decrease of 0.31%
    Rep 2022% = 24%     Rep 2020% = 24.74%   , percent change is -1.685578%
    Unaf 2022 = 20.2% Unaf 2020% = 18.7% , percent change is 9.4%
    
    There are a few interesting storylines here to review and one of them is that in September, two months out from election day, we saw the number of unafiliated voters rise which means people left their political affiliation. 

```{r}

2022 dems =  2250961
total = 4161637
percentage = (2250961/4161637)*100

2020 dems: 2258099
total = 4106456 
percentage= (2258099/4106456)*100

(2250961 - 2258099) / (2258099) *100

2022 Republican: 998905
total: 4161637
percentage: (998905/4161637)*100= 24%

2020 Republican: 1016031
total: 4106456
percentage= (1016031/4106456)*100
  
percent change: (998905-1016031)/(1016031) *100

2022 Unaf 841573
total:4161637
percentage (841573/4161637)*100

2020 Unaf 769201
total 4106456
percentage (769201/4106456)*100

percent change (841573-769201)/(769201)*100



```

Q2. Among the Maryland jurisdictions with the most registered voters, which one had the biggest percentage increase in unaffiliated voters between 2020 and 2022? How did that county vote for governor in 2018 and for president in 2020? Contrast the partisan percentages of that county with the county that had the second-largest percentage increase in unaffiliated voters. You should display information from the data that shows the counties with the largest percentage increase first.

A2. Out of the top few counties with the highest amount of total voters the county with the highest percent change in unaffiliated voters is PG county. PG county voted democrate in 2018 and democrat in 2020. Montgomery county had a more than double republican percentage in 2022.


```{r}
### this shows that Mongomery county has the most registered voters in 2022
current_registration_2022 %>% 
group_by(county) %>% 
select(county, unaf_17, total_19) %>% 
arrange(desc(total_19))

### totals for 2020.
current_registration_2020 %>% 
group_by(county) %>% 
select(county, unaf_23, total_25) %>% 
arrange(desc(total_25))


# Mutating Pr. George's to be the same


current_registration_2022 <- current_registration_2022 %>%
  mutate(county = case_when(county == "PR. GEORGE'S" ~ "PRINCE GEORGE'S",
                            TRUE ~ county))

# Left joining the data


registration_joined <- current_registration_2020 %>%
  left_join(current_registration_2022, by=c("county"))

registration_joined <- registration_joined %>%
  mutate(perc_change = (unaf_17-unaf_23)/(unaf_23)*100)

registration_joined %>% 
arrange(desc(total_19))

x <- registration_joined %>% 
  mutate(dem_perc = (dem_17/total_25)*100) %>%
  mutate(rep_perc = (rep_18/total_25)*100) %>% 
  mutate(unaf_perc =(unaf_17/total_25)*100)

```


Q3. Do the party percentages of new registrants (% Dem, % Rep, % Unaffiliated) from September 2022 and September 2020 match the overall percentage of partisan registration in the state at those times?

In other words, did new registrants mirror the overall partisan composition of the electorate, or were they different? Do any of the methods of registration vary significantly from the overall partisan pattern? Which method is worthy of further exploration?

A3. There are some changes in the party percentage. Republicans in Alleganey county saw a big drop from 2020-2022. I also noticed a big drop in the percentage of dems in Baltimore City. 

```{r}

### partisan breadkdown for new registatnts in 2022

perc_2022 <- new_party_registration_2022 %>% 
  mutate(dem_perc = (dem/total)*100) %>% 
  mutate(rep_perc = (rep/total)*100) %>% 
  mutate(unaf_perc = (unaf/total)*100) %>% 
select(county, rep_perc, dem_perc, unaf_perc)

#shows 2020
perc_2020 <- x %>% 
select(county, rep_perc, dem_perc, unaf_perc)

perc_total <- perc_2020 %>%
  left_join(perc_2022, by=c("county"))

perc_total %>% 
mutate(rep_difference = rep_perc.x-rep_perc.y) %>% 
mutate(dem_difference = dem_perc.x-dem_perc.y) %>% 
mutate(unaf_difference= unaf_perc.x-unaf_perc.y) %>% 
select(county, rep_difference, dem_difference, unaf_difference)






### I need desperate help with this one but I tried ###



```

Q4. Choose your own adventure! Using both the 2022 and 2020 data for one of the dataframes (changes, removals or new registrations), ask and answer a question that makes at least one comparison and requires code to answer. That could involve calculating a percentage, percentage change or the raw difference between two columns.

A4. 4.8 % of people changed their name in 2020 from PG county and 2.24% changed their name in 2022 from PG county.

```{r}

### What percentage changed their name in 2020 from PG county and what percentage changed their name in 2022 ###

party_changes_2020 %>% 
mutate(perc_names_2020 = (name/total))

party_changes_2022 %>% 
mutate(perc_names_2022 = (name/total)) %>% 
filter(county == "PR. GEORGE'S")




```

Q5. What's the best story idea or question you've seen as a result of the work you've done in this lab?

A5. One quetion I have that could be worth persuing is what could the rise in unafiliated voters be attributed to? What election news drew them away from their own political party and wanted to go unafiliated?