---
title: "state_profile"
author: "Dylan Manfre"
date: "2022-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("/Volumes/Elements/D-Drive/University of Maryland/Classes/Fall 2022/Data Journalism/Github Repository/Data_Journalism_2022_Fall/major_assignments/state_profile/")

```

```{r}
#libraries
library(tidyverse)
library(janitor)
library(lubridate)

```

**This section will be where I load in my datasets**
```{r}
# loading in the PA Democratic Contributions
pa_dem_contribs <- read_csv("data/pa_dem_contribs_21_22.csv")
head(pa_dem_contribs)

#laoding in GOP contribs
pa_gop_contribs <- read_csv("data/pa_gop_contribs_21_22.csv")
head(pa_gop_contribs)

#loading in Democratic Expenses
pa_dem_expenses <- read_csv("data/pa_dem_expenses_21_22.csv")
head(pa_dem_expenses)

#loading in GOP expenses
pa_gop_expenses <- read_csv("data/pa_gop_expenses_21_22.csv")
head(pa_gop_expenses)

```

**This section will be where I clean each dataset.**
**Each dataset will get its own codeblock for cleaning so that it is organized.**

```{r}
## Cleaning pa_dem_contribs


pa_dem_contribs <- pa_dem_contribs %>% 
select(-image_number, -line_number, -transaction_id, -file_number, -committee_name...9, -candidate_id, -candidate_prefix, -candidate_suffix, -candidate_office, -candidate_office_full, -candidate_office_state, -candidate_office_state_full, -candidate_office_district, -conduit_committee_id)

# renaming columns
pa_dem_contribs <- pa_dem_contribs %>% 
rename(committee_name = committee_name...2) %>% 
mutate(contribution_receipt_date = ymd(contribution_receipt_date)) %>% 
mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))


pa_gop_orgs

```

```{r}
# cleaning pa_dem_expenses

 pa_dem_expenses



```


```{r}
# cleaning pa_gop_contribs
pa_gop_contribs <- pa_gop_contribs %>% 
select(-image_number, -line_number, -transaction_id, -file_number, -committee_name...9, -candidate_id, -candidate_prefix, -candidate_suffix, -candidate_office, -candidate_office_full, -candidate_office_state, -candidate_office_state_full, -candidate_office_district, -conduit_committee_id) %>% 
rename(committee_name = committee_name...2) %>% 
mutate(contribution_receipt_date = ymd(contribution_receipt_date)) %>% 
mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))


```


```{r}
#cleaning pa_gop_expenses

pa_gop_expenses


```


***Story time***
In order to do a complete analysis of the Pennsylvania Republican and Democratic Parties, we need to understand and learn who the major players are. Basically, we need to answer the five W’s for each party: who, what, where, when and why. This is the basic level of analysis for the two parties but we truly can go through a rabbit hole in each file.

I’ll structure it like this: I’ll write a question, then below put the code block for the answer and then write the answer out with a piece of analysis under it. Let’s start with the Democratic contributions from Pennsylvania. Money in and money out are the two most important things in a financial report. Pennsylvania as of the analysis done Nov. 27, received $14,897,061. 

```{r}
# How much did Pennsylvania raise in total.. number as of Nov. 27

pa_dem_contribs %>% 
summarize(total_pa_dem_contribs = sum(contribution_receipt_amount))

# the result of this code is a 1x1 tibble with the total 14897061
```

It appears that Pennsylvania’s democratic party only got 15% of its overall donations from individuals which tells me that the average Joe in the state, is not pouring money into the committee. What surprised me about this was that individuals in the state gave $1,660,668 more than large political action committees did. 

```{r}
#showing how much each entity donated and breaking it down by percentage for dems
dem_entity_perc <- pa_dem_contribs %>%
group_by(entity_type, entity_type_desc) %>% 
summarize(total = sum(contribution_receipt_amount)) %>% 
mutate(perc_total = total/14897061*100) %>% 
arrange(desc(perc_total))

# showing difference between IND and PAC for Dems
ind_pac_dem_difference <- 2314694.2-654025.6
dem_entity_perc

#showing how much each entity donated and breaking it down by percentage for GOP
gop_entity_perc <- pa_gop_contribs %>%
group_by(entity_type, entity_type_desc) %>% 
summarize(total = sum(contribution_receipt_amount)) %>% 
mutate(perc_total = total/14897061*100) %>% 
arrange(desc(perc_total))

gop_entity_perc


```
Comparing the Dems to the GOP is interesting here as it shows that PACs still lags behind Individual contributors to the GOP which is noteworthy because I figured it would give more to the GOP. There is, however, a massive drop-off in the other committee category. It only makes up 2% of donations to the GOP where as the Democrats got 50% of its fundraising from that category alone.

We need to look at the top key players here. Who donated the most to Pennsylvania’s Democratic and Republican Committees and where did they come from? Are they prominent people at all? Did any one person make up a large percentage of the total donations? We’ll do this for the Republicans too.
```{r}
# Where did the GOP contribs come from.
pa_gop_contribs %>% 
group_by(contributor_name, contributor_employer) %>% 
summarize(state_total = sum(contribution_receipt_amount)) %>% 
mutate(perc_total = state_total/14897061*100) %>% 
arrange(desc(perc_total))
```
From here, we can use the same code to break this down by entity type to look at the specific individuals and analyze if they are prominent at all. 




```{r}
pa_gop_contribs %>% 
filter(entity_type == "IND") %>% 
group_by(contributor_name) %>% 
summarise(total = sum(contribution_receipt_amount)) %>% 
arrange(desc(total))
#Notes: 14 people gave the maximum of $20k in the period for the GOP


pa_dem_contribs %>% 
filter(entity_type == "IND") %>% 
group_by(contributor_name) %>% 
summarise(total = sum(contribution_receipt_amount)) %>% 
arrange(desc(total))
#Notes 7 people gave the max of 20k in the period for the dem campaign.


```




Here I will write an analysis of what these code blocks show and compare the top individual Republican and Democratic donors


**Put your analysis of the ORG findings here**

```{r}
# showing the orgs doners for DEMOCRATS
pa_dem_orgs <- pa_dem_contribs %>% 
filter(entity_type == "ORG") %>% 
group_by(contributor_name) %>% 
summarize(total = sum(contribution_receipt_amount)) %>% 
arrange(desc(total))

pa_dem_orgs

# showing the top doners for GOP
pa_gop_orgs <- pa_gop_contribs %>% 
filter(entity_type == "ORG") %>% 
group_by(contributor_name) %>% 
summarize(total = sum(contribution_receipt_amount)) %>% 
arrange(desc(total))

pa_gop_orgs

```










**Use this section for codes to work with before you put them in the story section **
```{r}
# showing the top donors, where they're from and how much they gave in total. 
pa_dem_contribs %>%
filter(entity_type == "IND") %>% 
group_by(contributor_name, contributor_city, contributor_state) %>% 
summarise(total_contrib = sum(contribution_receipt_amount)) %>% 
arrange(desc(total_contrib))

## State ranking and % from individuals not PACs or anything?
dem_state_perc_ind <- pa_dem_contribs %>%
filter(entity_type == "IND") %>% 
group_by(contributor_state) %>% 
summarize(total = sum(contribution_receipt_amount)) %>% 
mutate(perc_total = total/14897061*100) %>% 
arrange(desc(perc_total))
### The percentages from individuals is quite small here with PA's totals making up just 15%. This tells me that the Dem committee as a whole is not getting much money from individual donations and is likely getting more from others.

#Which entity gave the most?
dem_entity_perc <- pa_dem_contribs %>%
group_by(entity_type) %>% 
summarize(total = sum(contribution_receipt_amount)) %>% 
mutate(perc_total = total/14897061*100) %>% 
arrange(desc(perc_total))
### COM makes up 50% of the donations here which is "Other committees". 

#which organization gave the most to the dem party?
pa_dem_contribs %>% 
filter(entity_type == "ORG") %>% 
group_by(contributor_name) %>% 
count() %>% 
arrange(desc(n))
  
  
#summarize(total = sum(contribution_receipt_amount)) %>% 
count()


# Which contributor gave the most out of any category... Fetterman/Schumer are key names up there. 
pa_dem_contribs %>% 
filter(entity_type == "ORG") %>% 
group_by(contributor_name) %>% 
summarize(total = sum(contribution_receipt_amount)) %>% 
arrange(desc(total))



```








